### 2016-06-22 intro.tex
Notes

\section{Introduction}
We revisit the formalization of Bloom filters~\cite{xxx} and closely related data structures as primitives with explicit security goals.  Recently, Naor and Yogev~\cite{xxx} (hereafter NY) gave such a formalization, providing syntax for a generalized abstraction of Bloom filters, as well as casting the classical Bloom filter soundness condition as a security experiment admiting adaptive adversaries.  To briefly summarize, a Bloom filter is a data structure for compactly representing a set~$\mathcal{S}$ in a way that supports constant time responses to set-membership queries.  The representation of~$\mathcal{S}$ is an indexed array~$M$ of~$m$-bits, all initally set to 0.  To insert $x \in \mathcal{S}$ into~$M$, one computes indices $h_1(x),h_2(x),\ldots,h_k(x)$ where each $h_i \colon \mathcal{U} \to \{0,1,\ldots m-1\}$ maps from a universe~$\mathcal{U}$ (which contains $\mathcal{S}$) to indices of~$M$.  The bits at positions $h_1(x),h_2(x),\ldots,h_k(x)$ are set to 1; if a bit is already 1, due to inserting some previous element of~$\mathcal{S}$, then it is left alone.  After all elements of~$\mathcal{S}$ have been inserted, the resulting bit-array~$M$ is taken as the representation.  To test if $y \in \mathcal{S}$, one checks if  $(h_1(y)=1 \wedge h_2(y)=1 \wedge \cdots \wedge h_k(y)=1)$, returning ``yes'' if so, and ``no'' otherwise.   Bloom filters are correct, in the sense that if $y \in \mathcal{S}$, the set-membership test always returns ``yes''.  The classical soundness condition asserts that for all $y \notin\mathcal{S}$,  $\Prob{h_1,h_2,\ldots,h_k \getsr \mathrm{Func(\mathcal{U},\mathbb{Z}_m)}\colon h_1(y)=1 \wedge h_2(y)=1 \wedge \cdots \wedge h_k(y)=1} \leq \epsilon$ where $\epsilon$, the false-positive rate, is a parameter of the Bloom filter.  (The values of~$\epsilon,m,k,|\mathcal{S}|$ are related in a well known way.)  NY recast this soundness condition as a game in which an adversary~$A$ is given~$\mathcal{S}$ and an oracle for set-membership queries, and it attempts to find a false-positive.  The fact that~$A$ can \emph{adapt} its queries and its output based on what it learns from the set-membership oracle is the key definitional constribution of NY.

@@

Syntactically, NY says a Bloom filter (with a steady representation) is a pair of algorithms $\mathbf{B}=(\mathbf{B}_1, \mathbf{B}_2)$.  The setup algorithm~$\mathbf{B}_1$ is randomized, and on input a set~$\mathcal{S}$ it returns a representation~$M$ (from some undefined, implicit set of representations).  The query algorithm~$\mathcal{B}_2$ is determinisic, and on input a representation~$M$ and an element~$x \in \mathcal{U}$, it returns a bit.  For correctness they make the classical demand, that for all $x \in \mathcal{S}$ we have $\Prob{\mathbf{B}_2(\mathbf{B}_1(S),x)=1}=1$.  For soundness, adversary~$A$ plays the following game, ``Adversarial-resilient Bloom Filter with a steady representation''.  First, $M \getsr \mathbf{B}_1(\mathcal{S})$.  Then the adversary is given the set~$\mathcal{S}$ and may make queries $x_1,x_2,\ldots,x_q$ to an oracle for $\mathbf{B}_2(M,\cdot)$.  At the end of its execution, the adversary outputs~$x^*$ and wins the game if (1) $x^* \not\in \mathcal{S} \cup \{x_1,x_2,\ldots,x_q\}$, and (2) $\mathbf{B}_2(M,x^*)=1$.

First, we note that the security notion forbids~$A$ to win if it outputs $x^* \in \{x_1,x_2,\ldots,x_q\}$, even if one of these was a false-positive.  
%
Second, we note that the NY syntax does not capture the size of~$M$ as a parameter; the surrounding text does say that $\mathbf{B}_1$ ``outputs a compressed representation [of $\mathcal{S}$] $\mathbf{B}_1(\mathcal{S})=M$'' but what exactly this means is never addressed.  The formal syntax admits $\mathbf{B}_1(\mathcal{S})=\mathcal{S}$ and $\mathbf{B}_2(\mathcal{S},x)=1 \Leftrightarrow x \in \mathcal{S}$, which is also trivially secure by their definition.  These are both small, but important complaints.  But a larger complaint, and one that is much harder to see, is that that it is not clear that the NY formalism actually captures real world Bloom filters.

In particular, say that $\mathbf{B}_1$ is implementing a classic Bloom filter.  Then the hash functions $h_1,h_2,\ldots,h_k$ are either sampled from some implicit family~$\mathcal{H}$ by~$\mathbf{B}_1$ (recall that this algorithm is randomized), or they are hard-coded into the description of~$\mathbf{B}_1$.  Let's assume the former, for the moment.  In order for~$\mathbf{B}_2$ to be able to respond to set-membership queries, it must have $h_1,h_2,\ldots,h_k$ as well.  Since~$\mathbf{B}_2$ receives only the representation~$M$ as input, it must be the case that $h_1,h_2,\ldots,h_k$ are encoded into~$M$ somehow.  But the text strongly suggests, in several places, that~$M$ does \emph{not} include these.  Both NY and classical analysis of Bloom filters speak of the number of bits used to represent the set~$\mathcal{S}$: the classical analysis means only the number of bits in the actual array, and NY make comparisons throughout between the number of bits~$m$ used by their instantiations and the classical bounds.  Morevover, the Bloom filter they construct in one of their main positive results clearly \emph{does not} count as part of the representation size the hash functions used to indicate array locations. Thus, if we assume that $\mathbf{B}_1$ is sampling $h_1,h_2,\ldots,h_k$, it seems that their formalization fails to capture the basic Bloom filter construction; at best, it is hard to see clearly that it does.

On the other hand, we can assume that the hash functions are implicitly hard-coded into~$\mathbf{B}_1$ (and $\mathbf{B}_2$).  Then we can treat~$\mathbf{B}_1$ as deterministic, since it will simply carry out the steps required to populate an array using these hash functions, and output the array~$M$ as the representation.  If we apply Kirchoff's rule, the descriptions of~$\mathbf{B}_1,\mathbf{B}_2$ are known to the adversary, who is also provided with~$\mathcal{S}$.  In this case, the adversary can compute~$M$ itself, as well as test for false-positives itself.  So it isn't clear how the security notion helps to reason about difficulty of finding a false-positive.  Moreover, NY stresses (at least once) that the representation is \emph{not} given to the adversary in their notions.  So it seems unlikely that they intended for $h_1,h_2,\ldots,h_k$ to be hard-coded into the algorithms.

@@

We generalize this syntax to hash-based filters that are mutable, in the sense that they allow for updating of the representation.  This allows us to capture counting Bloom filters, spectral Bloom filters, stable Bloom filters, count-min sketches, and other compact set respresentation data structures...

>>>>

\jnote{Are we being too specific by focusing on hash-based
set-membership only? It would be nice to give general definitions
that apply to arbitrary data structures (constructed in arbitrary
ways) and then only later specialized to hash-based Bloom filters.}

\section{Introduction}
We revisit the formalization of Bloom filters~\cite{xxx} and closely
related data structures as primitives with explicit security goals.
Recently, Naor and Yogev~\cite{xxx} (hereafter NY) gave such a
formalization, providing syntax for a generalized abstraction of
Bloom filters, as well as casting the classical Bloom filter
soundness condition as a security experiment admiting adaptive
adversaries.  To briefly summarize, a Bloom filter is a data
structure for compactly representing a set~$\mathcal{S}$ in a way
that supports constant-time responses to set-membership queries. The
representation of~$\mathcal{S}$ is an bit-array~$M$ of length~$m$,
all initially set to~0.  To insert $x \in \mathcal{S}$ into~$M$, one
computes indices $h_1(x),h_2(x),\ldots,h_k(x)$ where each $h_i
\colon \mathcal{U} \to \{0,1,\ldots m-1\}$ maps from a
universe~$\mathcal{U}$ (which contains $\mathcal{S}$) to indices
of~$M$.  The bits at positions $h_1(x),h_2(x),\ldots,h_k(x)$ are set
to 1; if a bit is already 1, due to inserting some previous element
of~$\mathcal{S}$, then it is left alone.  After all elements
of~$\mathcal{S}$ have been inserted, the resulting bit-array~$M$ is
taken as the representation.  To test if $y \in \mathcal{S}$, one
checks if \[(M[h_1(y)]=1 \wedge M[h_2(y)]=1 \wedge \cdots \wedge
M[h_k(y)]=1),\] returning ``yes'' if so, and ``no'' otherwise. Bloom
filters are correct, in the sense that if $y \in \mathcal{S}$, the
set-membership test always returns ``yes''.  The classical soundness
condition asserts that for all $\mathcal{S}$ and any $y
\notin\mathcal{S}$, we have $\Prob{h_1,h_2,\ldots,h_k \getsr
\mathrm{Func}(\mathcal{U},[m])\colon M[h_1(y)]=1 \wedge \cdots
\wedge M[h_k(y)]=1} \leq \epsilon$ where $\epsilon$, the
false-positive rate, is a parameter of the Bloom filter.  (The
values of~$\epsilon,m,k$, and $|\mathcal{S}|$ are related in a
well-known way.) NY recast this soundness condition as a game in
which an adversary~$A$ is given~$\mathcal{S}$ and an oracle for
set-membership queries, and it attempts to find a false positive.
The fact that~$A$ can \emph{adapt} its queries and its output based
on what it learns from the set-membership oracle is the key
definitional distribution of~NY.

@@

Syntactically, NY says a Bloom filter (with a steady representation)
is a pair of algorithms $\mathbf{B}=(\mathbf{B}_1, \mathbf{B}_2)$.
The setup algorithm~$\mathbf{B}_1$ is randomized, and on input a
set~$\mathcal{S}$ it returns a representation~$M$ (from some
undefined, implicit set of representations).  The query
algorithm~$\mathcal{B}_2$ is deterministic, and on input a
representation~$M$ and an element~$x \in \mathcal{U}$, it returns a
bit.  For correctness they make the classical demand, that for all
$x \in \mathcal{S}$ we have
$\Prob{\mathbf{B}_2(\mathbf{B}_1(\mathcal{S}),x)=1}=1$.  For
soundness, adversary~$A$ plays the following game,
``Adversarial-resilient Bloom Filter with a steady representation''.
First, $M \getsr \mathbf{B}_1(\mathcal{S})$. \jnote{Where does the
set come from? Is it random? Adversarially chosen?} Then the
adversary is given the set~$\mathcal{S}$ and may make queries
$x_1,x_2,\ldots,x_q$ to an oracle for $\mathbf{B}_2(M,\cdot)$.  At
the end of its execution, the adversary outputs~$x^*$ and wins the
game if (1) $x^* \not\in \mathcal{S} \cup \{x_1,x_2,\ldots,x_q\}$,
and (2) $\mathbf{B}_2(M,x^*)=1$.

First, we note that the security notion forbids~$A$ to win if it
outputs $x^* \in \{x_1,x_2,\ldots,x_q\}$, even if one of these was a
false positive. \jnote{We discussed this on the phone, but let me
put it in writing here. I think this is actually a reasonable choice
because it captures the notion that ``even after finding some false
positives, it remains hard to find additional false positives.''
Note that in practice finding false positives is ``easy'' since
$\epsilon$ might be $1/1000$ rather than negligible.}
%
Second, we note that the NY syntax does not capture the size of~$M$
as a parameter \jnote{I think this is also reasonable; typically
security is defined without regard for efficiency}; the surrounding
text does say that $\mathbf{B}_1$ ``outputs a compressed
representation [of $\mathcal{S}$] $\mathbf{B}_1(\mathcal{S})=M$''
but what exactly this means is never addressed.  The formal syntax
admits $\mathbf{B}_1(\mathcal{S})=\mathcal{S}$ and
$\mathbf{B}_2(\mathcal{S},x)=1 \Leftrightarrow x \in \mathcal{S}$,
which is also trivially secure by their definition.  These are both
small, but important complaints.  But a larger complaint, and one
that is much harder to see, is that that it is not clear that the NY
formalism actually captures real world Bloom filters.

In particular, say we implement a classic Bloom filter.  Then the
hash functions $h_1, h_2, \ldots, h_k$ are either sampled from some
family~$\mathcal{H}$ by~$\mathbf{B}_1$ (recall that this algorithm
is randomized), or they are hard-coded into the description
of~$\mathbf{B}_1$.  Let's assume the former for the moment.  In
order for~$\mathbf{B}_2$ to be able to respond to set-membership
queries, it must have $h_1,h_2,\ldots,h_k$ as well.
Since~$\mathbf{B}_2$ receives only the representation~$M$ as input,
it must be the case that $h_1,h_2,\ldots,h_k$ are encoded into~$M$
somehow.  But the text strongly suggests, in several places,
that~$M$ does \emph{not} include these.  Both NY and classical
analysis of Bloom filters speak of the number of bits used to
represent the set~$\mathcal{S}$: the classical analysis means only
the number of bits in the actual array, and NY make comparisons
throughout between the number of bits~$m$ used by their
instantiations and the classical bounds. Moreover, the Bloom filter
they construct in one of their main positive results clearly
\emph{does not} count as part of the representation size the hash
functions used to indicate array locations.
%Thus, if we assume that
%$\mathbf{B}_1$ is sampling $h_1,h_2,\ldots,h_k$, it seems that their
%formalization fails to capture the basic Bloom filter construction;
%at best, it is hard to see clearly that it does.

\jnote{FWIW, I think the classical description of Bloom filters has
the same problem. I.e., if the hash functions are not fixed, then
their description should be counted as part of the length of the
data structure. On the other hand, if they are fixed, then the
false-positive rate is not well-defined (since there is no
probability space to speak of). So in that sense, NY just carried
over the problem from the classical literature.}

On the other hand, assume the hash functions are implicitly
hard-coded into~$\mathbf{B}_1$ (and $\mathbf{B}_2$). Then we can
treat~$\mathbf{B}_1$ as deterministic, since it will simply carry
out the steps required to populate an array using these hash
functions, and output the array~$M$ as the representation.  If the
descriptions of~$\mathbf{B}_1,\mathbf{B}_2$ are known to the
adversary, who is also provided with~$\mathcal{S}$, then the
adversary can compute~$M$ itself, as well as test for false
positives itself. So it isn't clear how the security notion helps to
reason about difficulty of finding a false-positive. Moreover, NY
stresses (at least once) that the representation is \emph{not} given
to the adversary in their notions.  So it seems unlikely that they
intended for $h_1,h_2,\ldots,h_k$ to be hard-coded into the
algorithms. \jnote{Perhaps it is reasonable to consider a definition
in which the set is not given to the adversary?}

@@

We generalize this syntax to hash-based filters that are mutable
\jnote{I find ``mutable'' a slightly awkward term,  maybe because it
makes me think of mutability in the context of programming
languages. In much data structures literature, this is called the
``dynamic'' case, in contrast to the ``static'' case.}, in the sense
that they allow for updating of the representation. This allows us
to capture counting Bloom filters, spectral Bloom filters, stable
Bloom filters, count-min sketches, and other compact set
representation data structures...
